Dependently Typed Multi-stage Calculus

### 2019-11-22 slides/main.tex
Spell checked.

\title{Dependently Typed Multi-stage Calculus}
\author{\underline{Akira Kawata} and Atsushi Igarashi}
\institute{Gradudate School of Informatics, Kyoto University}
\date{Dec 2, 2019}

@@

\begin{frame}[fragile]{Our Approach: \LMD}
    Introduce dependent types into \LTP [Hanada\&Igarashi'14].
    \begin{itemize}
        \item Features of multi-stage programming
        \item Features of dependent types
    \end{itemize}
    Our contribution
    \begin{itemize}
        \item Define typing, kinding, well-formed kind rules with stages
        \item Define equivalence rules including CSP
    \end{itemize}
    \note{
        Our approach to the problem of multi-stage programming is introducing dependent types into a multi-stage calculus by Hanada and Igarashi.
        We call our multi-stage calculus as \LMD.
    }
\end{frame}

@@

\begin{frame}[fragile]{Cross Stage Persistence (CSP)}
    Use normal values in code values
    \begin{tabbing}
        \hspace{5mm} \= \( \textbf{let}\ \text{mul3}\ x = 2 * x\ \textbf{in} \) \\
        \> \( \TB (\% \text{mul3}\ 5) \)
    \end{tabbing}
\end{frame}

@@

\begin{frame}[fragile]{Types of \LMD}
    \begin{block}{Types}
    \begin{tabbing}
        \hspace{5mm} \( \tau \) \= ::= X \hspace{20mm} \= (constants) \\
        \> \( \mid \TW\ \tau \) \> (code type) \\
        \> \( \bm{\mid \Pi x:\tau.\tau} \) \> (function dependent type) \\
        \> \( \bm{\mid \tau\ M} \) \> (application of term to dependent type)
    \end{tabbing}
    \end{block}
    We need CSP to use normal variables in code types.
    \begin{itemize}
        \item OK  \( \Pi n: \TW (\text{Vec}\ \%n) \)
        \item NG  \( \Pi n: \TW (\text{Vec}\ n) \)
    \end{itemize}
   \note{
    }
\end{frame}

@@

\begin{frame}[fragile]{Problem of type of vadd3}
    Hard to combine with other functions.
    \newcommand{\Vn}{\text{Vec}\ \%n}
    \newcommand{\Vpt}{\text{Vec}\ \%3}
    \newcommand{\Vt}{\text{Vec}\ 3}
    \begin{tabbing}
        \hspace{5mm} \= \( \text{val}\ \text{vadd3} : \TW (\Vpt \to \Vpt \to \Vpt) \) \\
        \> \( \text{val}\ \text{double} : \TW (\Vt \to \Vt) \) \\[2mm]
        \> \( \textbf{let}\ \text{add-double}\ v\ w = \) \\
        \> \hspace{5mm} \( \TB (\TBL\text{double}\ (\TBL \text{vadd3}\ \TBL v\ \TBL w));; \) \\
    \end{tabbing}
\end{frame}

@@

\begin{frame}[fragile]{Related Works}
    \begin{itemize}
        \item \LTP [Hanada\&Igarashi'14]
            \begin{itemize}
                \item Multi-stage programming including explicit CSP
            \end{itemize}
    \end{itemize}
    \note{
    }
\end{frame}

\begin{frame}[fragile]{Conclusion}
        We introduced dependent types into muti-stage programming.
        \begin{itemize}
            \item Define typing, kinding, well-formed kinding rules for multi-stage calculus.
            \item Define equality rules.
                \begin{itemize}
                        \item Especially, dealing with CSP.
                \end{itemize}
            \item Prove type soundness.
        \end{itemize}
    \note{
    }
\end{frame}

>>>>

\title{Dependently Typed Multi-stage Calculus}
\author{\underline{Akira Kawata} and Atsushi Igarashi}
\institute{Graduate School of Informatics, Kyoto University}
\date{Dec 2, 2019}

@@

\begin{frame}[fragile]{Our Approach: \LMD}
    Introduce dependent types into \LTP [Hanada\&Igarashi'14].
    \begin{itemize}
        \item Features of multi-stage programming
        \item Features of dependent types
    \end{itemize}
    Our contribution
    \begin{itemize}
        \item Define typing, kinding, well-formed kind rules for multi-stage calculus
        \item Define equivalence rules including CSP
    \end{itemize}
    \note{
        Our approach to the problem of multi-stage programming is introducing dependent types into a multi-stage calculus by Hanada and Igarashi.
        We call our multi-stage calculus as \LMD.
    }
\end{frame}

@@

\begin{frame}[fragile]{Cross Stage Persistence (CSP)}
    Enable using normal values in code values
    \begin{center}
        \( \textbf{let}\ \underbrace{\text{mul3}}_{\text{normal value}}\ x = 2 * x\ \textbf{in} \underbrace{\TB\ (\% \text{mul3}\ 5)}_{\text{code value}} \)
    \end{center}
    CSP symbol \( \% \) {\bf{lift up}} an normal value into code value.
\end{frame}

@@

\begin{frame}[fragile]{Types of \LMD}
    \begin{block}{Types}
    \begin{tabbing}
        \hspace{5mm} \( \tau \) \= ::= X \hspace{20mm} \= (constants) \\
        \> \( \mid \TW\ \tau \) \> (code type) \\
        \> \( \bm{\mid \Pi x:\tau.\tau} \) \> (function dependent type) \\
        \> \( \bm{\mid \tau\ M} \) \> (application, e.g. \(\text{Vec}\ 3\))
    \end{tabbing}
    \end{block}
    Note: We need CSP to use normal variables in code types.
    \begin{itemize}
        \item OK  \( \Pi n: \TW (\text{Vec}\ \%n) \)
        \item NG  \( \Pi n: \TW (\text{Vec}\ n) \)
    \end{itemize}
   \note{
    }
\end{frame}

@@

\begin{frame}[fragile]{Problem of type of vadd3}
    Hard to compose with other functions.
    \newcommand{\Vn}{\text{Vec}\ \%n}
    \newcommand{\Vpt}{\text{Vec}\ \%3}
    \newcommand{\Vt}{\text{Vec}\ 3}
    \begin{tabbing}
        \hspace{5mm} \= \( \text{val}\ \text{vadd3} : \TW (\Vpt \to \Vpt \to \Vpt) \) \\
        \> \( \text{val}\ \text{double} : \TW (\Vt \to \Vt) \) \\[2mm]
        \> \( \textbf{let}\ \text{add-double}\ v\ w = \) \\
        \> \hspace{5mm} \( \TB (\underbrace{\TBL\text{double}}_{\Vt \to \Vt}\ \underbrace{(\TBL \text{vadd3}\ \TBL v\ \TBL w)}_{\Vpt});; \) \\
    \end{tabbing}
    We cannot compose because of type mismatch.
\end{frame}

@@

\begin{frame}[fragile]{Related Works}
    \begin{itemize}
        \item \LTP [Hanada\&Igarashi'14]
            \begin{itemize}
                \item Multi-stage programming including explicit CSP
            \end{itemize}
        \item Concoqtion [Forgarty et al]
            \begin{itemize}
                \item Introduce indexed type into MetaOCaml.
            \end{itemize}
    \end{itemize}
    \note{
    }
\end{frame}

\begin{frame}[fragile]{Conclusion}
    We introduced dependent types into multi-stage programming.
    \begin{itemize}
        \item Define typing, kinding, well-formed kinding rules for multi-stage calculus.
            \item Define equality rules.
                \begin{itemize}
                        \item Especially, dealing with CSP.
                \end{itemize}
            \item Prove type soundness.
        \end{itemize}
    \note{
    }
\end{frame}
